\documentclass[conference]{IEEEtran}

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{enumitem}
\usepackage{hyperref}
\hypersetup{hidelinks}

\title{Tracks: Integer Programming Formulation and Resolution Approach}
\author{
\IEEEauthorblockN{CHACON Jose}
\IEEEauthorblockA{\textit{RO03, Games, Graphs and Operations Research}\\
\textit{ENSTA Paris}\\
Palaiseau, France}
\and
\IEEEauthorblockN{FLORIDO Santiago}
\IEEEauthorblockA{\textit{RO03, Games, Graphs and Operations Research}\\
\textit{ENSTA Paris}\\
Palaiseau, France}
}

\begin{document}
\pagenumbering{arabic}
\pagestyle{plain}
\maketitle
\thispagestyle{plain}

\begin{abstract}
This report presents a graph-based mixed-integer formulation and Python
implementation for the \textit{Tracks} logic puzzle. A rectangular puzzle grid
is modeled as an undirected grid graph, where cells are vertices and admissible
track continuities are edges. The puzzle is treated as a feasibility problem:
the objective is to find a single railway path between two prescribed terminals
while satisfying row and column clues, fixed hints, local degree rules, and
global connectedness. The formulation uses binary cell and edge variables
together with a single-commodity flow certificate that eliminates disconnected
cycles and extra components. The implementation translates this model into a
PuLP/CBC solver, an independent validator, constructive instance generation,
course-compatible workflow wrappers, dataset-solving tools, and a Pygame
interface for visualization and interactive experimentation. Computational
tests on generated instances from \(5\times5\) to \(16\times16\) illustrate
the complete modeling-to-code pipeline and the practical role of the flow
connectivity formulation.
\end{abstract}

\begin{IEEEkeywords}
graph theory, mixed-integer linear programming, feasibility problem, puzzle
solving, connectivity constraints, Pygame
\end{IEEEkeywords}

\input{sections/01_introduction.tex}
\input{sections/02_mathematical_model.tex}
\input{sections/03_methodology.tex}
\input{sections/04_results.tex}
\input{sections/05_conclusion.tex}

\begin{thebibliography}{00}
\bibitem{garey1979}
M. R. Garey and D. S. Johnson, \emph{Computers and Intractability: A Guide to
the Theory of NP-Completeness}. W. H. Freeman, 1979.

\bibitem{validi2022}
H. Validi, A. Buchanan, and E. Lykhovyd, ``Imposing contiguity constraints in
political districting models,'' \emph{Operations Research}, vol. 70, no. 2,
pp. 867--892, 2022, doi: 10.1287/opre.2021.2141.

\bibitem{miyazawa2021}
F. K. Miyazawa, P. F. S. Moura, M. J. Ota, and Y. Wakabayashi,
``Partitioning a graph into balanced connected classes: Formulations,
separation and experiments,'' \emph{European Journal of Operational Research},
vol. 293, no. 3, pp. 826--836, 2021, doi: 10.1016/j.ejor.2020.12.059.

\bibitem{ahuja1993}
R. K. Ahuja, T. L. Magnanti, and J. B. Orlin, \emph{Network Flows: Theory,
Algorithms, and Applications}. Prentice Hall, 1993.

\bibitem{python}
Python Software Foundation, ``Python Language Reference,'' 2026. [Online].
Available: \url{https://www.python.org}

\bibitem{pulp}
S. Mitchell, M. O'Sullivan, and I. Dunning, ``PuLP: A linear programming
toolkit for Python,'' 2011.
\end{thebibliography}

\end{document}
